---
title: "Data Visualisation 2019 - Assignment 4 -Mark Carley - 20071864"
output:
  html_notebook: default
  pdf_document: default
---


### List of R colors:

http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf



# The normal distribution

* The __normal distribution__ is the graph of the function
\[ f(x)=\frac{1}{\sqrt{2\pi \sigma}}e^{-\frac{(x-\mu)^2}{2\sigma}}\]
where $\mu$ (pronounced as "mu") is the __mean__ and $\sigma$ (pronounced as "sigma"") is the __standard deviaiton__ of the distribution.


## Example 1

* The __standard__ normal distribution is the one where $\mu=0$ and $\sigma=1$, and is shown in the graph below

```{r}
X<-seq(-5,5,len=201) #Create the x-axis between -5 and +5
P<-dnorm(X,mean=0,sd=1) # Evaluate the Normal Distribution function over the x-axis
plot(X,P,type='l',lwd=2,col='steelblue') # Plot the Probability along th x-axis

```
* The lines of this cell have the following functions:
    
1. __X<-seq(-5,5,len=200)__ This creates a  sequence of $x$-values, which will form the horizontal axis. These values start at -5, and stop at +5, and the sequence has 200 points altogether.
    
```{r}
X  
```
    
2. __P<-dnorm(X,mean=0,sd=1)__ This evaluates the normal distribution with __mean=0__ and __standard deviation = 1__. With __mean=0__ this means the peak of the curve is at $x=0$. The $sd=1$ command determines how spread out the graph is, the larger __sd__ the more spread out the graph.
   
3. The last line instruct __R__ to plot __X__ on the horizontal axis and __P__ on the vertical axis. The argument __type='l'__ instructs __R__ to plot the curve as a line, __lwd=2__ indicates a __linewidth__ of 2 and __col='steelblue'__  indicates the __colour__ to plot the curve. 


## Exercise 1

Plot the normal disribution with $\mu=4$ and $\sigma=2.15$. Note that you will have to re-define an appropriate __X__-interval for this plot also. A convient way to do this so that your plot looks symmetric is to set 
\[\mathbf{X=\text{seq}(\mu-5*\sigma,\mu+5*\sigma)}\]
and __do not__ forget to include * when you want to denote multiplication.

# Interpretation of the normal curve
* When $x$ is __normally distributed__ with mean $\mu$ and standard deviation $\sigma$, then the __proability__ that $x$ lies between the values $x_1$ and $x_2$ is the __area__ under the normal distribution curve, between these two values.

* An improtant feature of every normal distribution curve is the the __total area__ beneath the curve is always equal to 1.



## Exercise 1 solutions

----------------------------------------------------------------------------------------------------------------------


```{r}
X<-seq(-5,5,len=201) 
P<-dnorm(X,mean=4,sd=2.15) 
plot(X,P,type='l',lwd=2,col='red') 
```

----------------------------------------------------------------------------------------------------------------------


## Example 2
Illustrate the probability for $x$ in the intervavl $-2\leq x\leq 1$ in the normal distribution with $\mu=0$ and $\sigma=1$. Calculate this proability.

```{r}
# Part 1. Plot the normal curve again - See Example 1
X<-seq(-5,5,len=201)
P<-dnorm(X,mean=0,sd=1)
plot(X,P,type='l',lwd=2,col='steelblue')

# Part 2. Shade the required area.
X1=seq(-2,1,length=201) # The interval -5<x<5
P1=dnorm(X1,mean=0,sd=1)
polygon(c(-2,X1,1),c(0,P1,0),col=adjustcolor("cornflowerblue",alpha.f=0.3),border=NA)
```

* The code in the cell above has the following meaning:

1. __Part 1__ This code has alredy been explained in __Example 1__
    
2. __Part 2__ We create a new __x1__ axis between __-5__ and __-1__. We create the corresponding __y-values__ using the __dnorm()__ funciton applied to the intrval __x1__. __polygon(c(-5,x_1,-1),c(0,y1,0)) indicates that the area between __-5__ and __-1__ on the __x__-axis and between __0__ and the corresponging values of __y1__ should be filled in.      
    
3. __col=adjustcolor("cornflowerblue",alpha.f=0.3)__ indicates the fill color and the transparencey __alpha__ of the filling. The smaller the value of __alpha__ the more transparent the filling, where __alpha__ should always be between __0__ and __1__.


# Probabilities
* The probability that $x$ lies between $-\infty$ and $-1$ is found using the function __dnorm()__

```{r}
dnorm(-1,mean=0,sd=1)
```

* The probability that $x$ lies between $-\infty$ and $2$ is

```{r}
dnorm(2,mean=0,sd=1)
```

* The the probability that __x__ lies between __-1__ and __2__ is 
```{r}
dnorm(2,mean=0,sd=1)-dnorm(-1,mean=0,sd=1)
```

* This correspond to the shaded area shown above.

# Confidence Intervals

Given a population, a __Confidence Interval (CI)__ is a means of estimating a __population statistic__ from the corresponding __sample staistic__ with a certain level of confidence. For example, we may take survey data on a collection of people to estimate the average mothly salary of the people in the survey. Then based on this sample average we can estimate a range for the average slalary of all people in the population, to a certain confidence level. The range of possible values for the population average is called a confidence interval.

## Example 3: 

* Suppose we wish to determine the mean salary of the population of a country with a certain degree of confidence.

* In practice it would be difficult, if not impossible, to collect data on the salary of everybody from the population.

* Instead, we may collect salary data from numerous samples of people from the population, and then get the mean salary for each of these samples.

* For each sample, the __sample mean__ would be spread with a normal distribution, whose standard deviation and mean would be close to the mean of the overall population.

* We let $\bar{x}$ denote the mean salary from one of these samples, and the standard deviation of the sample be $\sigma$ (this measures how much the means are spread about the true mean).

* We can now say that the true population mean lies in the interval
\[\bar{x}\pm z_{\frac{\alpha}{2}}\frac{\sigma}{\sqrt{n}},\]
where $n$ is the size of the sample for which the mean $\bar{x}$ was calculated.

* The parameter $\alpha$ is called the __confidence parameter__ associated with the confidence level, and the number $z_{\frac{\alpha}{2}}$ is the __z-score__ associated with the parameter $\frac{\alpha}{2}$.

* From a sample of 100 people, we find that the lowest salary from the sample is EUR 15000 and the highest is EUR 55000, with a mean salary of EUR 35000 and a standard deviation of EUR 2500.

* We can plot the normal distribution for this data as follows:

  1. We create a sample size between 15000 and 55000 (the sample range), with step size 50. We call this sample data __Sample1__
  
  2. We use the __dnorm()__ function to indicate that the data in __Sample1__ is __normally distributed about the mean__ where the mean is 35000 and the standard deviation is 2500.
  
  3. We plot the distribution using the usual __plot()__ function, with __Sample1__ on the x-axis and __Data1__ on the y-axis.
  
```{r}
Sample1<-seq(15000,55000, by=50)
Data1 <- dnorm(Sample1,mean=35000, sd=2500)
plot(Sample1,Data1,ylab="Rel. Freq.",xlab="Salaries (EUR)",type="l",lwd=2,col="red")
```
## A 90% Confidence Interval

* Suppose we wish to determine the __average__ salary value at a confidence level of  __90%__.

* This would correspond to the salary interval which gives 90% of the area beneath the normal distribution curve above.

* This number can be calculated using a normal distribution table, and finding the $z_{\alpha}$-score corresponding to $\alpha=1-0.90=0.1$.

```{r}
# Part 1. Plot the normal curve again - See Example 1
Z0<-seq(-5,5,len=200)
H0<-dnorm(Z0,mean=0,sd=1)
plot(Z0,H0,type='l',lwd=2,col='steelblue',main='The 90% confidence interval')

# Part 2. Shade the required area.
x1=seq(-qnorm(0.95),qnorm(0.95),length=200)
y1=dnorm(x1,mean=0,sd=1)
polygon(c(-qnorm(0.95),x1,qnorm(0.95)),c(0,y1,0),col=adjustcolor("cornflowerblue",alpha.f=0.3),border=NA)
```

* When we look for a confidence interval of 90% we are looking for the values of $z$ such that 90% of the area beneath the curve, between $-z_{\frac{\alpha}{2}}$ and $z_{\frac{\alpha}{2}}$.

* It also means we are looking for the $z$-values which will exclude 5% of the area in each tail, that is we are looking for the $z$-values for each of the dashed line in the plot below: 


```{r}
Standard<-seq(-4,4,by=0.001)
StandardData<-dnorm(Standard,mean=0,sd=1)
plot(Standard,StandardData,ylab="Rel. Freq.", xlab="z", type="l", lwd=2,col="red")
segments(-qnorm(0.9),0,-qnorm(0.9),dnorm(-qnorm(0.9),mean=0,sd=1),lwd=2.5,lty=2)
segments(qnorm(0.9),0,qnorm(0.9),dnorm(qnorm(0.9),mean=0,sd=1),lwd=2.5,lty=2)

```
* 90% of the area beneath the curve is contained __between__ the two dashed lines.

* 5% of the area is in each tail to the left and right.

* It also means that the area beneath the curve containing the left tail and the middle portion is 95%

* To find the $z$-value corresponding to this 95 % in __*R*__ we can use the __qnorm()__ function as follows:
```{r}
z90=qnorm(0.95)
z90
```


* In the plot above this means that 95% of the are beneath the curve is between $z=-\infty$ and $1.64485$. 

* It also means that  __90%__ of the area lies, from $z=-1.644854$ to $z=-1.644854$.

* For the purposes of our salary survey above, it also means 
\[z_{\frac{\alpha}{2}}=1.644845\]

* In terms of our survey we are also 90% certain that the __mean population salry__ lies in the range

```{r}
lower_bound=35000-z90*2500/sqrt(100)
upper_bound=35000+z90*2500/sqrt(100)
lower_bound
upper_bound
```
### We are 90% confident that the avareage salary of the overall population lies in the range €34,588.79 to €35,411.21.

In terms of our normal distribution plot this means, we are 90% certain the __mean population salary__ lies between the two dashed blue lines.
```{r}
Sample1<-seq(15000,55000, by=50)
Data1 <- dnorm(Sample1,mean=35000, sd=2500)
plot(Sample1,Data1,ylab="Rel. Freq.",xlab="Salaries (EUR)",type="l",lwd=2,col="red")
abline(v=lower_bound,lwd=2,lty=2,col="blue")
abline(v=upper_bound,lwd=2,lty=2,col="blue")
```


## Exercise 2

A company wishes to determine the cost associated with providing health insurance for all of its employees. The company asked a sample of 35 employees how much they currently pay annually for health insurance. The data obtained revealed a sample mean premium of EUR635 with a standard deviation of EUR67. Using this data answer the following

1. Find the 90%, 95%, 96% and 99% confidence intervals for the average amount spent by all customers.
2. Plot a normal distribution for this data set showing the confidence intervals obtained.


## Exercise 2 solutions

---------------------------------------------------------------------------------------------------------------------

1. 

```{r}
xbar2 <-635
sd2<- 67
n2<-35
SE2<-sd2/sqrt(n2)
```

## 90% confidence interval


```{r}
x90<-qnorm(0.95)
x90
```

## Lower Bound 90

```{r}
CI90_L<-xbar2-x90*SE2
CI90_L
```

## Upper Bound 90

```{r}
CI90_H<-xbar2+x90*SE2
CI90_H
```

## 95% confidence interval

```{r}
x95<-qnorm(0.975)
x95
```

## Lower Bound 95

```{r}
CI95_L<-xbar2-x95*SE2
CI95_L
```

## Upper Bound 95

```{r}
CI95_H<-xbar2+x95*SE2
CI95_H
```

## 96% confidence interval

```{r}
x96<-qnorm(0.985)
x96
```

## Lower Bound 96

```{r}
CI96_L<-xbar2-x96*SE2
CI96_L
```

## Upper Bound 96

```{r}
CI96_H<-xbar2+x96*SE2
CI96_H
```

## 99% confidence interval

```{r}
x99<-qnorm(0.995)
x99
```

## Lower Bound 99

```{r}
CI99_L<-xbar2-x99*SE2
CI99_L
```

## Upper Bound 99

```{r}
CI99_H<-xbar2+x99*SE2
CI99_H
```


---------------------------------------------------------------------------------------------------------------------


### Plotting

```{r}
xaxis<-seq(xbar2-4*sd2,xbar2+4*sd2,by=2)
xaxis
```

```{r}
ydata<-dnorm(xaxis,mean=xbar2,sd=sd2)
ydata
```

```{r}
plot(xaxis,ydata,lwd=2,type='l',col='red')
abline(v=CI90_L,lty=1,lwd=2,col='blue')
abline(v=CI90_H,lty=1,lwd=2,col='blue')

abline(v=CI95_L,lty=1,lwd=2,col='orange')
abline(v=CI95_H,lty=1,lwd=2,col='orange')

abline(v=CI96_L,lty=1,lwd=2,col='green')
abline(v=CI96_H,lty=1,lwd=2,col='green')

abline(v=CI99_L,lty=1,lwd=2,col='grey')
abline(v=CI99_H,lty=1,lwd=2,col='grey')
```
blue = 90 , orange = 95 , green = 96 , grey = 99



## Exercise 3

The mean number of emails received by employees at a company was measured on a ample of 58 employees. The sample mean was found to be 21 emails per day with a standard deviation of 9. Using this data answer the following

1. Find the 90%, 94%, 96% and 99% confidence intervals for the average amount spent by all customers.
2. Plot a normal distribution for this data set showing the confidence intervals obtained.


## Exercise 3 Solutions

---------------------------------------------------------------------------------------------------------------------

1.

```{r}
xbar3 <-21
sd3<- 9
n3<-58
SE3<-sd3/sqrt(n3)
```

## Lower Bound 90

```{r}
CI901_L<-xbar3-x90*SE3
CI901_L
```

## Upper Bound 90

```{r}
CI901_H<-xbar3+x90*SE3
CI901_H
```

## Lower Bound 95

```{r}
CI952_L<-xbar3-x95*SE3
CI952_L
```

## Upper Bound 95

```{r}
CI952_H<-xbar3+x95*SE3
CI952_H
```

## Lower Bound 96

```{r}
CI963_L<-xbar3-x96*SE3
CI963_L
```

## Upper Bound 96

```{r}
CI963_H<-xbar3+x96*SE3
CI963_H
```

## Lower Bound 99

```{r}
CI994_L<-xbar3-x99*SE3
CI994_L
```

## Upper Bound 99

```{r}
CI994_H<-xbar3+x99*SE3
CI994_H
```


---------------------------------------------------------------------------------------------------------------------

2.


```{r}
xaxis2<-seq(xbar3-4*sd3,xbar3+4*sd3,by=2)
xaxis2
```

```{r}
ydata2<-dnorm(xaxis2,mean=xbar3,sd=sd3)
ydata2
```

```{r}
plot(xaxis2,ydata2,lwd=2,type='l',col='red')
abline(v=CI901_L,lty=1,lwd=2,col='blue')
abline(v=CI901_H,lty=1,lwd=2,col='blue')

abline(v=CI952_L,lty=1,lwd=2,col='orange')
abline(v=CI952_H,lty=1,lwd=2,col='orange')

abline(v=CI963_L,lty=1,lwd=2,col='green')
abline(v=CI963_H,lty=1,lwd=2,col='green')

abline(v=CI994_L,lty=1,lwd=2,col='grey')
abline(v=CI994_H,lty=1,lwd=2,col='grey')
```
blue = 90 , orange = 95 , green = 96 , grey = 99



# Samples, Means and Standard Deviations


## The Mean

Given a data set we can calculate the mean and the standard deviation of the data-set using the functions __mean()__ and __sd()__. 


Given a data set of size $n$, with data values $x_1,x_2,\ldots,x_n$, the mean of the data set is given by
\[
\bar{x}=\frac{\displaystyle{\sum_{i=1}^{n}}x_{i}}{N}.
\]
This can be evaluated in __R__ using the function __mean()__:

## Example 2

Suppose the data from a survey is given by the data set
\[
1.3,-4.2,1.55,6.78,4.5,3.21, 5.89,7.34,-6.88,-7.9
\]
Create a data structure in __R__ to represent this data set. Use the __mean()__ function to evaluate the mean of the data set.

```{r}
dataset2<-c(1.3,-4.2,1.55,6.78,4.5,3.21, 5.89,7.34,-6.88,-7.9)
mean(dataset2)
```
* Hence the mean is
\[\bar{x}=1.159\]

## The Standard Deviation

Given a data set we can calculate the mean and the standard deviation of the data-set using the functions __mean()__ and __sd()__.

Given a data set of size $n$, with data values $x_1,x_2,\ldots,x_n$, the mean of the data set is given by
\[
s=\frac{\displaystyle{\sum_{i=1}^{n}}(x_{i}-\bar{x})^2}{N-1}.
\]
This can be evaluated in __R__ using the function __mean()__:

## Example 4

Evaluate the standard deviation of the data set in __Example 2__ using the __sd()__ function.
```{r}
sd(dataset2)
```

* Hence the standard deviation is 
\[
s= 5.608865
\]


## Example 5
The inflation rate in Ireland between the years 1992 and 2016 is given in the data file __IrelandInflation.csv__(_Source:_ [World Bank](http://databank.worldbank.org/data/home.aspx)).  We will import this into __R__ using the usual command __read.csv(file.choose())__ as a data structure. Using this data structure we want to answer the following:

1. Find the mean inflation rate between the years 1992 and 2016
2. Find the standard deviation of this data set
3. Find the number of data points in this data set
4. Using these three values, find the range of possible values for the average inflation rate in Ireland over all years with 95% confidence
5. Plot the inflation data as a normal distribution.


### Importing the Data

```{r}
InflationData<-read.csv("IrelandInflation.csv")
InflationData
```
### Creating the Data Structure
```{r}
Inflation<-InflationData$Inflation.Rate..
Inflation
```
### Part 1 - The Mean Inflation

```{r}
xbar=mean(Inflation)
xbar
```


* The mean inflation rate between 1992 and 2016 in Ireland was $\bar{x}=2.076174%$


### Part 2 - The Standard Deviation 

```{r}
s=sd(Inflation)
s
```
* The standard deviation  in the rate of inflation between 1992 and 2016 is $\sigma=2.189926$

### Part - The Number of Data Points

We use the __length()__ function to determine the number of data-points in __Inflation__
```{r}
n=length(Inflation)
n
```
* There are $n=25$ data points in the data set __Inflation__

### Part 4 - The 90% Confidence Interval
* The confidence parameter $\alpha$ is given by
\[\alpha = 1-0.95=0.05\Rightarrow \frac{\alpha}{2}=0.025\]

* We now wish to find the $z_{\frac{\alpha}{2}}$ which gives an area of $0.95+0.025=0.975$ under the curve of the normal distribution. We find this $z$-score using __qnorm()__

```{r}
z0=qnorm(0.975)
z0
```


* For the 95% confidence interval, we have $z_{\frac{\alpha}{2}}=1.959964$. Hence the upper and lower bound for the 95% confidence interval for the data set are

```{r}
xlower=xbar-z0*s/sqrt(n)
xupper=xbar+z0*s/sqrt(n)

xlower
xupper
```

#### We are 95% certain that the mean inflation rate for Ireland lies in the interval 1.217739% to 2.934609%.


### Part 5 - The Normal Distribution Plot

We now plot __StandardInflation__ to give the normal distribution plot:
  
  1. First we create an interval appropriate  for creating a normal distribution form the data-set __Inflation__:
    For this plot, we will make the plot range start at $\bar{x}-5\sigma$  and end at $\bar{x}+5\sigma$, with a step-size      0.0001:  
```{r}
Interval1=seq(xbar-5*s,xbar+5*s,by=0.0001)
```
  2. Next we create a normal distribution from this data with mean and standard deviation evaluated earlier:
```{r}
NormalDistInflation=dnorm(Interval1, mean=xbar, sd=s)
``` 
  3 Now we plot __NormalDistInflation__ on the y-axis and __Interval1__ on the x-axis:
```{r}
plot(Interval1,NormalDistInflation,ylab="Rel. Freq.",xlab="Inflation Rate %",type="l",lwd=2,col="red")
abline(v=xlower,lwd=2,lty=2,col="blue")
abline(v=xupper,lwd=2,lty=2,col="blue")
```
## Exercise 4: 

* The data file __IrelandEmissions.csv__ contains the CO2 emissions, in metric tons per capital, of Ireland between 1992 and 2016.
(_Source_: [World Bank](http://databank.worldbank.org/data/home.aspx))

* Using the data in this file, create a data structure to capture the numerical data over those 25 years. 

* Using this data structure answer the following:

1. Find the mean inflation rate between the years 1992 and 2016
2. Find the standard deviation of this data set
3. Find the number of data points in this data set
4. Using these three values, find the 90% and 99% confidence intervals for the average CO2 emissions of Ireland over all years.
5. Plot the inflation data as a normal distribution, and indicate the confidence intervals on these plots.

---------------------------------------------------------------------------------------------------------------------

1.

```{r}
IrelandEmissions<-read.csv("IrelandEmissions.csv")
IrelandEmissions
```

```{r}
IrelandEmissions<-IrelandEmissions$CO2.Emissions..
IrelandEmissions
```

```{r}
xbar=mean(IrelandEmissions)
xbar
```

The mean inflation rate is 9.541227.

---------------------------------------------------------------------------------------------------------------------

2.

```{r}
s=sd(IrelandEmissions)
s
```

Standard deviatioh is 1.192195.

---------------------------------------------------------------------------------------------------------------------

3.

```{r}
n=length(IrelandEmissions)
n
```

Number of data points = 23.

---------------------------------------------------------------------------------------------------------------------

4.

```{r}
xbar4 <-23
sd4<- 1.192
n4<-9.54
SE4<-sd4/sqrt(n4)
```

## Lower Bound 90

```{r}
CI902_L<-xbar4-x90*SE4
CI902_L
```

## Upper Bound 90

```{r}
CI902_H<-xbar4+x90*SE4
CI902_H
```

## Lower Bound 99

```{r}
CI9942_L<-xbar4-x99*SE4
CI9942_L
```

## Upper Bound 99

```{r}
CI9942_H<-xbar4+x99*SE4
CI9942_H
```

---------------------------------------------------------------------------------------------------------------------

5.

```{r}
xaxis3<-seq(xbar4-4*sd4,xbar4+4*sd4,by=2)
xaxis3
```

```{r}
ydata3<-dnorm(xaxis3,mean=xbar4,sd=sd4)
ydata3
```

```{r}
plot(xaxis2,ydata2,lwd=2,type='l',col='red')
abline(v=CI902_L,lty=1,lwd=2,col='blue')
abline(v=CI902_H,lty=1,lwd=2,col='blue')

abline(v=CI9942_L,lty=1,lwd=2,col='grey')
abline(v=CI9942_H,lty=1,lwd=2,col='grey')
```




# Interpretation of Confidence Intervals
* We record some data associated with a __*population*__, given in the data set __X__ below:
```{r}
X=c(0.34,0.77,0.98,1.03,1.45,1.87,1.99,2.11,2.23,2.24,2.87,2.89,2.97,3.04,3.45,3.56,3.77,3.79,3.81,3.97,4.12,4.34,4.88,5.01,5.16)
```

* Obviously, the __population mean__ is easily available to us in this case:
```{r}
mean(X)
```

## Sampling the population

* In many cases of practical interest, this mean is not available to us, and the best we can do is form a __confidence interval__ for this mean.

* To form these confidence intervals, we take __samples__ of the population data.

* This is implemented in __R__ as follows:
```{r}
n=4 # sample size
N=10 # number of samples
Sample<-replicate(N,sample(X,n))
```
* We have taken __10 Samples__ with each sample  having __4 data values__

```{r}
Sample
```

* Now we wish to take the __mean of each sample__, which we do as follows:
```{r}
M <- numeric(length = length(1:N))
for (i in 1:N) {
  M[i] <- mean(Sample[,i])
}
M
```

* Clearly, the mean of each sample does not match the mean of the population.


## Confidence Intervals from the Samples

* We now form the __60% Confidence Interval__ associated with each sample mean, and to do so we need the __z score__ corresponding to the 80th percentile:

```{r}
z<-qnorm(0.80,mean=0,sd=1)
z
```

* Next we form the __standard error__ associated with the population and the samples:

1. The __standard deviation__ of the population is 
```{r}
sigma<-sqrt(sum((X-mean(X))^2)/length(X))
sigma
```
We do this because the function __sd()__ uses the formula for sample standard deviation.

2. The __standard error__ is 
```{r}
SE<-sigma/sqrt(n)
SE
```
i.e. the population standard deviation divided by the sample size.


3. The __lower boundaries__ of the confidence intervals are now:
```{r}
ML<-M-z*SE
ML
```

The __upper ends__ of the confidence intervals are:
```{r}
MU<-M+z*SE
MU
```

* The 60% confidence intervals following from these 10 samples are:

```{r}
paste(ML,MU,sep=' to ')
```

__Recall:__ The actual mean of the pupulation is

```{r}
mean(X)
```

* We deliberitely chose 60% confidence intervals to highlight the point that not all confidence intervals contain the true population mean.

* This gives us the correct interpretation for the 60% confidence interval:
 
    * A confidence iterval either does or does not contain the true population mean.
 
    * As we can see, if we take enough samples, then some of these confidence intervals will contain the mean and some will not.
 
    * If we take a large enough number of samples, then on average, 60% of the confidence intervals deduced from those samples, at confidence level 60%, will contain the true mean.
 
 
## Exercise 5:

1. Using the same data set __X__, take __N=50__ samples of sample size __n=5__.

2. Find the mean of each sample.

3. Find the __90%__ confidence interval associated with these sample averages.

4. Do all these confidence intervals contain the true population mean?


---------------------------------------------------------------------------------------------------------------------

1. 

```{r}
X=c(0.34,0.77,0.98,1.03,1.45,1.87,1.99,2.11,2.23,2.24,2.87,2.89,2.97,3.04,3.45,3.56,3.77,3.79,3.81,3.97,4.12,4.34,4.88,5.01,5.16)
```

```{r}
n=5 
N=50 
Sample<-replicate(N,sample(X,n))
```

```{r}
Sample
```

---------------------------------------------------------------------------------------------------------------------

2.

```{r}
MeanX <- numeric(length = length(1:N))
for (i in 1:N) {
  MeanX[i] <- mean(Sample[,i])
}
MeanX
```

---------------------------------------------------------------------------------------------------------------------

3.

```{r}
Xinterval<-qnorm(0.975,mean=0,sd=1)
Xinterval
```

---------------------------------------------------------------------------------------------------------------------







