---
title: "Data Visualisation 2019 - Assignment 9"
output: html_notebook
---

####  List of R colours: http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

# Cluster Analysis - Part II
    
### Example 1 - Continued from Assignment 8

1. Create __12__ data vectors to represent each of the following.
\[
\underline{a}=(3986,1943,444)\\
\underline{b}=(4271,1444,344)\\
\underline{c}=(8443,2020,415)\\
\underline{d}=(1898,1877,311)\\
\underline{e}=(2916,1253,289)\\
\underline{f}=(7476,1583,347)\\
\underline{g}=(3690,2131,491)\\
\underline{h}=(1355,1837,265)\\
\underline{i}=(7251,1492,306)\\
\underline{j}=(6498,1683,405)\\
\underline{k}=(5211,1298,367)\\
\underline{l}=(8001,1950,392)\\
\]

2. Use the __R__-function __rbind()__  to create a __matrix__ representing these vectors in a single data structure.  

3.  Use the __scale()__-function, so that all vector dimensions are of similar size.

4. Create a __scatter plot__ of the data in dimensions 1 and 2 of this scaled data structure.

5. Using this scatter plot, identify any __potential__ clusters among the 12 vectors.

### Solution

__Part 1.__ We begin by creating a list of data structures corresponding to the data objects given.

```{r}
a<-c(3986,1943,444)
b<-c(4271,1444,344)
c<-c(8443,2020,415)
d<-c(1898,1877,311)
e<-c(2916,1253,289)
f<-c(7476,1583,347)
g<-c(3690,2131,491)
h<-c(1355,1837,265)
i<-c(7251,1492,306)
j<-c(6498,1683,405)
k<-c(5211,1298,367)
l<-c(8001,1950,392)
```

__Part 2.__ The __single__ data structure is given by:

```{r}
Data<-rbind(a,b,c,d,e,f,g,h,i,j,k,l)
Data
```

__Part 3.__ The __scaled__ data structure is given by:

```{r}
Data_S<-scale(Data,center=F,scale=c(1000,1000,100))
Data_S
```



__Part 4.__  The __scatter plot__ of the scaled data structure in dimensions 1 and 2 is given by:

```{r}
plot(Data_S[,1],Data_S[,2],col='steelblue',pch=15,xlab='Scaled Dimension 1',ylab='Scaled Dimension 2',main='Scatter Plot for Data in Dimensions 1 and 2')
text(x=Data_S[,1]+0.2,y=Data_S[,2],labels=c('a','b','c','d','e','f','g','h','i','j','k','l'),col='darkslategray')
```

__Part 5.__ 
It appears from this scatter plot that the data is clustered as
\[(a,g), (b,k), (c,l), (d,h), e, (f,i,j).\]

__Remark:__ 
  * While this clustering appears natural when we use the scatter plot, we must also consider that we are only comparing the data vectors along __two__   dimensions. 
  
  * When we compare the data vectors along another two dimensions we __may__ get a completely different clustering.
  
  * When we perform a full cluster analysis, we take into account all dimensions of the objects in the data-set.


### Exercise 1

Eight makes of car were compared using three different criteria:


  1. Price (euro)    2. Engine (cc)    3. Efficiency (km/L)

The data collected are given in the table below:

| Make       | Price (euro) | Engine (cc) | Efficiency |
|------------|-----------|-------------|------------|
| Audi       | 38812     |     1968    |     22.7   |
| BMW        | 35571     |     1995    |     23.7   |
| Citroen    | 20451     |     1560    |     24.7   |
| Hyundai    | 23620     |     1685    |     23.7   |
| Jaguar     | 53693     |     1999    |     26.5   |
| Mercedes   | 41909     |     1950    |     25.5   |
| Mitsubishi | 28192     |     2268    |     18.8   |
| Toyota     | 27978     |     1995    |     21.6   |

Using the data in this table, answer the following:

1. Create 8 __data vectors__ to represent each car make.

2. Combine these data vectors using the __rbind()__ function, to create a single __data structure__.

3. Rescale the dimensions of this data stricture, so each dimension is measured in the same order.

4. Create a scatter-plot of this data along dimensions __1__ and __2__ of this data set.

5. Use this scatter plot to identify potential clusters in the data-set.

6. Create a scatter-plot of this data along dimensions __1__ and __3__ of this data set.

7. Use this scatter plot to identify potential clusters in the data-set.

8. Do the clusters match?

----------------------------------------------------------------------------------------------------------------------

1.

```{r}
audi<-c(38812,1968,22.7)
bmw<-c(35571,1995,23.7)
citreon<-c(20451,1560,24.7)
hyundai<-c(23620,1685,23.7)
jaguar<-c(53693,1999,26.5)
mercedes<-c(41909,1950,25.5)
mitsubishi<-c(28192,2268,18.8)
toyota<-c(27978,1995,21.6)
```

2.

```{r}
Data2<-rbind(audi,bmw,citreon,hyundai,jaguar,mercedes,mitsubishi,toyota)
Data2
```

3.

```{r}
Scale_cars<-scale(Data2,center=F,scale=c(10,10,10))
Scale_cars
```

4.

```{r}
plot(Scale_cars[,1],Scale_cars[,2],col='red',pch=15,xlab='Scaled Dimension 1',ylab='Sclaed Dimension 2',main='Scatter Plot for Data for Dimensions 1 and 2')
text(x=Scale_cars[,1]+0.2,y=Scale_cars[,2],labels=c('audi','bmw','citreon','hyundai','jaguar','mercedes','mitsubishi','toyota'),col='darkslategray')
```

5.

From the data set, the data appears to be clusteres as (hyundai, citreon), (toyota,mitsubishi), (bmw,audi) or (audi,mercedes).


6.

```{r}
plot(Scale_cars[,1],Scale_cars[,3],col='red',pch=15,xlab='Scaled Dimension 1',ylab='Sclaed Dimension 2',main='Scatter Plot for Data for Dimensions 1 and 2')
text(x=Scale_cars[,1]+0.2,y=Scale_cars[,3],labels=c('audi','bmw','citreon','hyundai','jaguar','mercedes','mitsubishi','toyota'),col='darkslategray')
```

7.

From this scatter plot the data appears to be clustered as (citreon,hyundai), (toyota,mitsubishi) and (bmw,audi).

8.

No the clusters do not match.




## Clustering Algorithms

* To implement a clustering algorithm on a set of data vectors, we must find the distances between those data vectors, using a metric such as the Euclidean or Manhattan metric.

* Once we have these distances we can use a clustering algorithm such as __Complete Linkage__ or __Single Linkage__ to identify the clusters in the data-set.

### Example 2:

Using the scaled data-frame from __Example 1__, answer the following:

  1. Find the __Manhattan__ distances between these data vectors.
  
  2. Using these Manhattan distances, use the function __hclust()__ to perform a cluster analysis on this data using the __Single Linkage__ method.
  
  3. Use the function __fvis_dend()__ to create a __dendrogram__ for the data with 2,4,6 and 8 clusters.
  
_Solution:_

__Part 1:__ The Manhattan distances between the scaled data vectors are given by

```{r}
Dist_S_M=dist(Data_S,method='manhattan')
Dist_S_M
```
  
__Part 2:__ The clustering is performed using the function __hclust()__, which is available from the library __cluster__.

* Importing the library:
```{r}
library(cluster)
```
  
* Performing the clustering using __Single Linkage__ 

```{r}
Cluster_1_M=hclust(Dist_S_M,method='single')
Cluster_1_M
```
  
__Part 3:__ The function __fviz_dend()__ is available in the __factoextra__ library, which we import as follows:
```{r}
library(factoextra)
```

* The cluster dendrogram with __2 clusters__ is given by
```{r}
fviz_dend(Cluster_1_M,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

* The cluster dendrogram with __4 clusters__ is given by
```{r}
fviz_dend(Cluster_1_M,k=4,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue",'seagreen4','goldenrod'))
```

* The cluster dendrogram with __6 clusters__ is given by
```{r}
fviz_dend(Cluster_1_M,k=6,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue",'seagreen4','goldenrod','azure4','coral'))
```

* The cluster dendrogram with __8 clusters__ is given by
```{r}
fviz_dend(Cluster_1_M,k=8,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue",'seagreen4','goldenrod','azure4','coral', 'purple', 'firebrick'))
```


### Exercise 2:
Using the  scaled car-data vectors created in in __Exercise 1__, answer the following:

  1. Find the __Manhattan__ distances between these data vectors.
  
  2. Using these Manhattan distances, use the function __hclust()__ to perform a cluster analysis on this data using the __Complete Linkage__ method.
  
  3. Use the function __fvis_dend()__ to create a __dendrogram__ for the data with 2,4,6 and 8 clusters.

----------------------------------------------------------------------------------------------------------------------

1.

```{r}
Dist_S_Cars=dist(Scale_cars,method='manhattan')
Dist_S_Cars
```

2.

```{r}
library(cluster)
```
```{r}
Cluster_1_Cars=hclust(Dist_S_Cars,method='complete')
Cluster_1_Cars
```

3.

```{r}
library(factoextra)
```

```{r}
fviz_dend(Cluster_1_Cars,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_Cars,k=4,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue",'seagreen4','goldenrod'))
```

```{r}
fviz_dend(Cluster_1_Cars,k=6,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue",'seagreen4','goldenrod','azure4','coral'))
```

```{r}
fviz_dend(Cluster_1_Cars,k=8,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue",'seagreen4','goldenrod','azure4','coral', 'purple', 'firebrick'))

```


### Exercise 3:

Six makes of laptop we compared using three different criteria:

  1. Storage (GB) 2. Screen (inches)  3. Ram (GB) 4. Clock-speed (GHz)
  
The data is available at 

  __Moodle__ $\rightarrow$ __Data Visualisation__ $\rightarrow$ __Data Files__ $\rightarrow$ __LaptopData__
  
* Import this data using __read.csv()__ in the usual way.

```{r}
Laptops<-read.csv(file.choose())
Laptops
```



* Modify this data structure so that it can used by the function __dist()__ (see Assignment 8).

_Step 1_
```{r}
LaptopNames<-as.character(Laptops$Make)
LaptopNames
```    
_Step 2_
```{r}
LaptopData<-as.data.frame(Laptops[,-1],row.names=LaptopNames)
LaptopData
```

Using this data structure answer the following:

 1. Rescale this new data structure so that all dimensions have values between 0 and 10.
 
 2. Find the Euclidean and Manhattan Distances between these scaled vectors.
 
 3. Apply the clustering algorithm using __Complete Linkage__ to both of these distance sets.
 
 4. Plot a __dedrogram__ with 2, 3, 4 and 5 data clusters obtained using each metric.
 
 
 ---------------------------------------------------------------------------------------------------------------------
 
 1.
 
```{r}
laptops1<-read.csv('LaptopData.csv')
laptops1
```
 
```{r}
LaptopNames<-as.character(Laptops$Make)
LaptopNames
```    

```{r}
LaptopData<-as.data.frame(Laptops[,-1],row.names=LaptopNames)
LaptopData
```

```{r}
dell<-c(1000,15.6,8,1.8)
acer<-c(500,15.6,4,1.7)
lg<-c(1000,15.6,16,1.8)
lenovo<-c(512,13.9,16,1.8)
google<-c(512,12.3,16,1.3)
delli3<-c(500,11.6,4,1.6)
```

```{r}
DataBind<-rbind(dell,acer,lg,lenovo,google,delli3)
DataBind
```



```{r}
Data_ScaleLaptops<-scale(DataBind,center=F,scale=c(125,5,4,1))
Data_ScaleLaptops
```

2.

```{r}
Dist_EL<- dist(Data_ScaleLaptops,method="euclidean")
Dist_EL
```

```{r}
Dist_ML<- dist(Data_ScaleLaptops,method="manhattan")
Dist_ML
```

3.

```{r}
Cluster_1_EL=hclust(Dist_EL,method='complete')
Cluster_1_EL
```

```{r}
Cluster_1_ML=hclust(Dist_ML,method='complete')
Cluster_1_ML
```

4.

```{r}
fviz_dend(Cluster_1_EL,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL,k=3,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL,k=4,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL,k=5,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```



### Exercise 4:

On Moodle, the file __EuroZoneData2017.csv__  compares the countries of the Euro Zone (excluding Malta as not all data was available), using 4 different criteria

  1. Population 2. GDP per capita (US\$) 3. Total Exports (US\$) 4. Total Imports (US\$)

__Moodle__ $\rightarrow$ __Data Visualisation__ $\rightarrow$ __Workbook Files__ $\rightarrow$ __EuroZoneData2017.csv__ 

(_Source:_ http://www.worldbank.org)

Using this data answer the following:

1. Import the data into this __R__ workbook using __read.csv()__.

2. Create an appropriate data frame from this data file which can be used by the __dist()__ function.

3. Rescale the data columns of this new data frame so that all dimensions have a similar size.

4. Find the Manhattan distances using this re-scaled data frame.

5. Apply the clustering algorithms __Complete Linkage__ and __Single Linkage__ to both of these distances.
 
6. Plot a __dedrogram__ with 2, 10, 15 and 20 data clusters obtained using both clustering methods. 

----------------------------------------------------------------------------------------------------------------------

1.

```{r}
eurozone<-read.csv('EuroZoneData2017.csv')
eurozone
```

2.

```{r}
eurozone1<-as.character(eurozone$Country)
eurozone1
```

```{r}
eurozoneData<-as.data.frame(eurozone[,-1],row.names=eurozone1)
eurozoneData
```

```{r}
dist(eurozoneData, method='euclidean')
```

3.

```{r}
Scaled_Eurozone_Data <- scale(eurozoneData,center=F,scale=c(100000,1000,10000000,1000000000))
Scaled_Eurozone_Data
```

4.

```{r}
Eurozone_Distances_M<-dist(Scaled_Eurozone_Data,method='manhattan')
Eurozone_Distances_M
```

```{r}
Eurozone_Distances_E<-dist(Scaled_Eurozone_Data,method='euclidean')
Eurozone_Distances_E
```


5.

```{r}
Cluster_1_EL1=hclust(Eurozone_Distances_M,method='complete')
Cluster_1_EL1
```

```{r}
Cluster_1_ML1=hclust(Eurozone_Distances_E,method='complete')
Cluster_1_ML1
```

```{r}
Cluster_1_EL2=hclust(Eurozone_Distances_M,method='single')
Cluster_1_EL2
```

```{r}
Cluster_1_ML2=hclust(Eurozone_Distances_E,method='single')
Cluster_1_ML2
```


6.

```{r}
fviz_dend(Cluster_1_EL1,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL1,k=10,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL1,k=15,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL1,k=20,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```




```{r}
fviz_dend(Cluster_1_ML1,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_ML1,k=10,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_ML1,k=15,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_ML1,k=20,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```




```{r}
fviz_dend(Cluster_1_EL2,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL2,k=10,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL2,k=15,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_EL2,k=20,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```




```{r}
fviz_dend(Cluster_1_ML2,k=2,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_ML2,k=10,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_ML2,k=15,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```

```{r}
fviz_dend(Cluster_1_ML2,k=20,color_labels_by_k=T, rect=T, k_colors=c("red", "dodgerblue"))
```





